\section{Cost and efficiency}

\begin{definition}[\textit{Pareto efficient}]
    An equilibrium is Pareto efficient if it is not possible to increase the utility of a player without decreasing the utility of some other player. 
\end{definition}
\noindent Nash equilibria need to be Pareto efficient and in fact they can be bad for all the players. 
To quantify how bad an outcome is, lut us consider costs, rather than utilities. 

The quality of a strategy profile $x=(x_1,\dots,x_n)$ is measured by a social cost function $x\mapsto C(x)$ from $X$ to $\mathbb{R}_{+}$.
The smaller $C(x)$ the better the outcome $x$. 
\begin{definition}[\textit{Benchmark}]
    The benchmark is the minimal value that a benevolent social planner could achieve: 
    \[\text{opt}=\min_{x\in X}C(x)\]
\end{definition}
\noindent For $x\in X$ the ration $\frac{C(x)}{\text{opt}}$ measures how far outcome $x$ is from being optimal. 
So, a large value implies a big loss in terms of social welfare, whereas a quotient close to 1 implies that $x$ is almost as efficient as an optimal solution. 

\begin{definition}[\textit{Price of anarchy}]
    Let $NE\subseteq X$ be the set of pure Nash equilibria of a cost game. 
    The price of anarchy is defined as: 
    \[\text{PoS}=\max_{\bar{x}\in NE}\dfrac{C(\bar{x})}{\text{opt}}\]
\end{definition}
\begin{definition}[\textit{Price of stability}]
    Let $NE\subseteq X$ be the set of pure Nash equilibria of a cost game. 
    The price of stability is defined as: 
    \[\text{PoS}=\min_{\bar{x}\in NE}\dfrac{C(\bar{x})}{\text{opt}}\]
\end{definition}
Note that $1\leq\text{PoS}\leq\text{PoA}$: 
\begin{itemize}
    \item $\text{PoA}\leq\alpha$ means that in every possible pure equilibrium the social cost $C(\bar{x})$ is no worse than $\alpha$ opt. 
    \item $\text{PoS}\leq\alpha$ means that there exists some equilibrium with social cost at most $\alpha$ opt. 
\end{itemize}

\begin{proposition}
    Consider a cost minimization finite potential game with potential $p:X\rightarrow\mathbb{R}$, and suppose that there exists $\alpha,\beta\geq 0$ such that: 
    \[\dfrac{1}{\alpha}C(x)\leq p(x)\leq \beta C(x)\qquad \forall x \in X\]
    Then, $\text{PoS}\leq \alpha\beta$
\end{proposition}
\begin{proof}
    Let $\bar{x}$ be a minimum of $p(\cdot)$ so that $\bar{x}$ is a Nash equilibrium. 
    For all $x \in X$: 
    \[\dfrac{1}{\alpha}C(\bar{x})\leq p(\bar{x})\leq p(x)\leq \beta C(x)\qquad \forall x \in X\]
    Since this is true for all $x$, we can choose the strategy $x$ yielding the optimal outcome $\text{opt}=\min_{x\in X} C(x)$, and hence it follows that $C(\bar{x})\leq \alpha\beta\text{opt}$. 
\end{proof}
\noindent In case a game deals with utilities rather than costs, one defined: 
\[\text{opt}=\max_{x\in X}U(x)\]
Here, $U(x)$ is some fixed utility function.