\section{Exercise one}

A small manufacturing company, known for being one of the most important producers of a specialized musical instrument, falls victim to a ransomware attack. 
This attack, initiated by malware designed to encrypt all files on the infected computer until a ransom is paid to the attacker, quickly spreads to all computers in use by the company.
\begin{enumerate}
    \item Identify and describe the two most critical threats/risks in this scenario, including the at-risk asset and suggesting one or two possible countermeasures for each
    \item Identify the possible threat agents based on the risks identified in point one.
\end{enumerate}

\subsection*{Solution}
\begin{enumerate}
    \item The two most critical threats/risks in this scenario are:
    \begin{itemize}
        \item Loss of business-critical data: this involves the potential loss of vital data such as key intellectual property, which could hinder the company's ability to continue producing its specialized goods.
            \begin{itemize}
                \item Asset at risk: business-critical data.
                \item Countermeasure: implement regular backups of all important data to ensure data recovery in case of an attack.
            \end{itemize}
        \item Loss of production time: the ransomware attack could result in significant downtime required to restore infected computers and systems, leading to a halt in production.
            \begin{itemize}
                \item Asset at risk: company's production capabilities.
                \item Countermeasure: implement redundant systems, isolate critical systems, and establish procedures for rapid disaster recovery to minimize downtime and economic losses.
            \end{itemize}
    \end{itemize}
    \item The possible threat agents in this scenario are:
        \begin{itemize}
            \item Cybercriminals: motivated by the potential for ransom payments, cybercriminals deploy ransomware attacks to extort money from victims.
            \item Competitors: a competitor may seek to damage the company's business operations or cause financial harm by disrupting production through a ransomware attack.
            \item Malicious traders: if the victim company is publicly listed, a malicious trader may exploit the resulting stock market decline caused by the ransomware attack for personal gain.
        \end{itemize}
\end{enumerate}