\section{Exercise 3}

Consider an Internet-connected smart speaker equipped with a voice-controlled intelligent virtual assistant, installed within a residence.
The speaker is linked to a wireless network and connected to a cloud service account.
It operates by continuously monitoring for a specific keyword. 
Upon detection of the keyword, the device records a brief audio clip, which is then uploaded to a cloud-based speech recognition service.
Subsequently, the device executes the requested action as per the recognized command.
These actions may include searching for specific information on the internet or interacting with the owner's cloud account.
Additionally, the device functions as a home automation hub, allowing voice commands to control various smart devices such as lights, door locks, heating systems, and more.
\begin{enumerate}
    \item Identify the most valuable assets at risk in this scenario.
    \item Suggest at least two potential attack surfaces of this smart speaker.
    \item Provide, in rough order of prevalence, the most likely potential digital attacks in this scenario.
\end{enumerate}

\subsection*{Solution}
\begin{enumerate}
    \item The most valuable assets at risk in this scenario include:
        \begin{itemize}
            \item Personal information such as musical preferences and location.
            \item Owners' voice, which is recorded for commands and has the potential to capture unintended conversations due to the device's always-listening microphone.
            \item The security of the actual house, particularly with the possibility of remotely controlling the door.
            \item The reputation of the device vendor.
        \end{itemize}
    \item The potential attack surfaces of this smart speaker encompass:
        \begin{itemize}
            \item The voice command interface.
            \item The cloud backend, susceptible to exploits or data breaches.
            \item The local network.
            \item Physical access to the device.
        \end{itemize}
    \item The most likely potential digital attacks in this scenario involve:
        \begin{itemize}
            \item Compromising the cloud vendor to access recordings, user data, and potentially gain control of the house.
            \item Malicious voice commands issued by a physical person or via a recording, such as a deceptive TV advertisement or malware playing a command to exploit the virtual assistant.
            \item Compromising the device from the local network to access information or monitor the user.
        \end{itemize}
\end{enumerate}