\section{Formal language theory}

A formal language is composed of words formed by selecting letters from an alphabet, and these words must adhere to a defined set of rules to be considered well-structured.
\begin{definition}
    An \emph{alphabet} $\Sigma$ is a finite collection of elements referred to as \emph{characters}, denoted as $\{ a_1,a_2,\dots, a_k \}$. 

    The \emph{cardinality} of an alphabet $\Sigma =\{ a_1,a_2,\dots, a_k \}$ represents the number of characters it encompasses, denoted as $\left\lvert \Sigma \right\rvert = k$. 

    A \emph{string} is a sequential arrangement of elements from the alphabet, potentially with repetitions.
\end{definition}
\begin{example}
    The alphabet $\Sigma =\{ a,b \}$ consists of two distinct characters.
    From this alphabet, various languages can be generated, including:
    \begin{itemize}
        \item $L_1=\{aa,aaa\}$
        \item $L_2=\{aba,aab\}$
        \item $L_3=\{ab,ba,aabb,abab,\dots,aaabbb,\dots\}$
    \end{itemize}
    In these languages, different combinations of the alphabet's characters are used to form words.
\end{example}
\begin{definition}
    The strings of a language are called \emph{sentences} or \emph{phrases}. 
    
    The \emph{cardinality} of a language is the number of sentence it contains.
\end{definition}
\begin{example}
    Considering the language $L_2=\{ bc,bbc \}$, it is evident that its cardinality is two.
\end{example}
\newpage
\begin{definition}
    The count of times a specific letter appears in a word is referred to as the \emph{number of occurrences}. 
    
    The \emph{length} of a string corresponds to the total number of elements it contains.

    Two strings are considered  \emph{equal} if and only if the following conditions are met:
    \begin{itemize}
        \item They possess the same length.
        \item Their elements match from left to right, sequentially.
    \end{itemize}
\end{definition}
\begin{example}
    In the string $aab$, the number of occurrences of the letters $a$ and $c$ is denoted as follows:
    \[{\left\lvert aab \right\rvert}_a = 2\]
    \[{\left\lvert aab \right\rvert}_c = 0\]
    The length of the string $aab$ is determined as:
    \[\left\lvert aab \right\rvert = 3\]
\end{example}