\section{Dependability principles}

Dependability is a critical consideration both during the design phase and during runtime operations.
During the design phase, it is essential to:
\begin{itemize}
    \item Analyze the system under development.
    \item Evaluate and measure its dependability properties.
    \item Make necessary modifications to the design as needed.
\end{itemize}
During runtime, the focus shifts to:
\begin{itemize}
    \item Detecting any malfunctions or failures that occur.
    \item Investigating and understanding the root causes of these issues.
    \item Taking appropriate reactive measures to address and mitigate the impact of the malfunctions.
\end{itemize}
Failures are commonplace in both the development and operational stages: while those in development should be averted, operational failures, being unavoidable due to the nature of system components, must be managed effectively. 
Design processes should factor in these potential failures to ensure that control and safety measures remain intact even when failures arise. 
Moreover, the effects of these failures should be predictable and deterministic rather than catastrophic.

Once upon a time, dependability was primarily a concern in safety-critical and mission-critical application domains like space exploration, nuclear facilities, and avionics. 
This was largely due to the significant cost associated with ensuring dependability, which was deemed acceptable only when absolutely necessary.
In non-critical systems, operational failures can lead to economic losses and damage to reputation, as seen in consumer products. 
However, in mission-critical systems such as satellites, automatic weather stations, surveillance drones, and unmanned vehicles, operational failures can result in serious or irreversible consequences for the mission at hand.
Safety-critical systems, on the other hand, pose a direct threat to human life if they fail during operation.
Examples include aircraft control systems, medical instrumentation, railway signaling, and nuclear reactor control systems.