\section{Technology}

\begin{definition}[\textit{Technology}]
    A technology represents a process that a given organization can perform, together with all the resources needed to perform the process. 
\end{definition}
\begin{definition}[\textit{Techincal system}]
    A technical system represents a set of machines supporting a given technology.
\end{definition}
\begin{definition}[\textit{Information system}]
    An information system is a set of coordinated processes producing an information output and executing information processing activities. 
\end{definition}
\begin{definition}[\textit{Information technology architecture}]
    An information system is a technology, an IT architecture is a techncial system supporting a given information system.
\end{definition}

\noindent For a long time, there has been an ongoing debate about how technical innovation influences organizations. 
A well-established set of beliefs links technological advancements to organizational change, shaping how companies adapt and evolve:
\begin{enumerate}
    \item \textit{Efficiency over effectiveness}: technological innovation primarily enhances efficiency rather than improving overall effectiveness. 
        It streamlines processes but doesn't necessarily guarantee better decision-making or outcomes.
    \item \textit{Economies of scale}: as technology advances, businesses can scale operations more efficiently, reducing costs per unit as production increases.
    \item \textit{Larger optimal size}: the minimum viable size of an organization tends to grow with technological progress, as larger entities can better leverage new systems.
    \item \textit{Increased specialization}: automation and sophisticated systems often lead to a workforce that is more specialized, with employees focusing on narrower, highly technical roles.
    \item \textit{Tayloristic perspective}: the traditional view, inspired by Taylorism, assumes that an optimal organizational structure exists.
    \item \textit{Limited focus on group work}: early studies largely ignored the impact of technology on teamwork and collaboration, focusing instead on individual efficiency.
    \item \textit{Greater bureaucracy and formalization}: as technical systems evolve, so do organizational rules, procedures, and levels of bureaucracy, making work more structured but also more rigid.
    \item \textit{More complex management}: with increased technology comes greater managerial complexity, requiring leaders to navigate intricate systems, regulations, and workflows.
\end{enumerate}