\section{Regular expressions}

Text documents are fundamentally just sequences of characters. 
Regular expressions provide a powerful way to search within these sequences by defining patterns that match specific character sequences.
Regular expressions are useful for: 
\begin{itemize}
    \item \textit{Pattern detection}: determine whether a specific pattern exists within a document.
    \item \textit{Information extraction}: locate and extract relevant information from a document whenever the pattern appears.
\end{itemize}
\renewcommand{\arraystretch}{1.25}
\begin{table}[!ht]
    \centering
    \begin{tabular}{|l|l|p{9.5cm}|}
    \hline
    \textbf{Name} & \textbf{Formula} & \textbf{Description} \\
    \hline
    Exact match & \texttt{aaa} & Matches the exact sequence \texttt{aaa} \\
    \hline
    Sequence choice & \texttt{(aaa|bbb)} & Matches either \texttt{aaa} or \texttt{bbb} \\
    \hline
    Wildcard & \texttt{.} & Matches any single character except for a newline \\
    \hline
    Character choice & \texttt{[]} & Matches any one character inside the square brackets \\
    \hline
    Newline & \textbackslash \texttt{n} & Represents a newline character \\
    \hline
    Tab & \textbackslash \texttt{t} & Represents a tab character \\
    \hline
    Whitespace & \textbackslash \texttt{s} & Matches any whitespace character \\
    \hline
    Non-whitespace & \textbackslash \texttt{S} & Matches any non-whitespace character \\
    \hline
    Digit & \textbackslash \texttt{d} & Matches any digit (\texttt{[0-9]}). \\
    \hline
    Word character & \textbackslash \texttt{w} & Matches any word character (\texttt{[a-zA-Z0-9]}) \\
    \hline
    Zero or more times & \texttt{*} & Matches the preceding character zero or more times \\
    \hline
    One or more times & \texttt{+} & Matches the preceding character one or more times \\
    \hline
    Zero or one time & \texttt{?} & Matches the preceding character zero or one time \\
    \hline
    Exactly $n$ times & \{\texttt{n}\} & Matches $n$ occurrences of the preceding character \\
    \hline
    From $n$ to $m$ times & \{\texttt{n,m}\} & Matches between $n$ and $m$ occurrences of the preceding character\\
    \hline
    \end{tabular}
\end{table}
\renewcommand{\arraystretch}{1}
    
\subsection{Regular expressions in text mining}
Regular expressions offer a powerful way to define patterns that can extract specific content from text documents. 
This allows for highly customizable and efficient text processing.

\noindent The advantages of regular expression based text extraction are: 
\begin{itemize}
    \item \textit{Simplicity}: regular expressions are a straightforward way to specify patterns.
    \item \textit{Precision}: extraction rules can be finely tuned to target specific patterns, which reduces false positives.
\end{itemize}
\noindent The limitations of regular expression based text extraction are: 
\begin{itemize}
    \item \textit{Manual rule creation}: writing extraction rules usually requires manual effort, which can be time-consuming and complex.
    \item \textit{False positives}: regular expressions can still yield false positives, where the pattern matches unintended content. 
    \item \textit{False negatives}: false negatives occur when the rules are not broad enough to capture all valid cases.
    \item \textit{Lack of context awareness}: regular expressions typically work on isolated patterns, without understanding the context in which the pattern appears.
\end{itemize}