\section{The process and product}

The creation of a software program necessitates a well-defined process. 
Both the software itself and the procedures employed possess inherent quality, and it is imperative for the software engineer to strive for the highest quality possible. 
This is essential because the process directly influences the ultimate outcome.

Software distinguishes itself from conventional product types in several ways, as it is:
\begin{enumerate}
    \item \textit{Intangible}, making it challenging to precisely describe and evaluate.
    \item \textit{Adaptable or malleable}, capable of being modified to suit evolving requirements.
    \item \textit{Human-intensive}, devoid of straightforward manufacturing processes.
\end{enumerate}
Software quality is influenced by various critical factors, including development technology, process quality, the caliber of individuals involved, cost considerations, and adherence to predefined schedules. 
Software quality attributes encompass:
\begin{itemize}
    \item \textit{Correctness}: ensuring that the software aligns with specified requirements.
    \item \textit{Reliability}: minimizing the probability of failure absence over a defined period.
    \item \textit{Robustness}: assessing the software's ability to perform reasonably in unforeseen scenarios.
    \item \textit{Performance}: evaluating the efficient utilization of resources.
    \item \textit{Usability}: focusing on the ease of use for the anticipated user base.
    \item \textit{Maintainability}: the capacity to facilitate software maintenance.
    \item \textit{Reusability}: extending maintainability to components, promoting their reuse.
    \item \textit{Portability}: the adaptability of software to diverse target environments.
    \item \textit{Interoperability}: ensuring harmonious coexistence and cooperation with other applications.
\end{itemize}
Process quality attributes consist of:
\begin{itemize}
    \item \textit{Productivity}: assessing the efficiency and output of the development process.
    \item \textit{Unity of effort} (person-month): measuring the collective contribution of effort over a month.
    \item \textit{Delivered item} (lines of code and function points): quantifying the tangible output of the development process.
    \item \textit{Timeliness}: the capability to respond to change requests in a timely fashion.
\end{itemize}