\section{Dart}

Dart is an object-oriented programming language developed by Google, designed to be easy to learn and highly efficient for building applications across various platforms. 
Key features of Dart include:
\begin{itemize}
    \item \textit{Object-oriented design}: Dart is a class-based language with single inheritance, utilizing a familiar C-style syntax that makes it accessible to developers with backgrounds in languages like Java or C++.
    \item \textit{Rich type system}: Dart supports interfaces, abstract classes, generics, and both optional and strong typing, providing flexibility while ensuring type safety.
    \item \textit{Everything is an object}: in Dart, everything is treated as an object, including numbers, functions, and even null values.
        This uniformity simplifies programming and enhances code consistency.
    \item \textit{Optional type annotations}: while Dart supports type annotations, they are optional, allowing developers to choose how explicitly they want to define types.
    \item \textit{Function and variable declarations}: Dart allows for the declaration of top-level functions as well as functions associated with classes. 
        Similarly, it supports both top-level variables and class-bound variables.
    \item \textit{Private identifiers}: any identifier that starts with an underscore is considered private to its library, encapsulating functionality and enhancing modularity.
\end{itemize}
These features make Dart a powerful choice for developers working with Flutter, enabling them to create robust and efficient applications.