\section{Introduction}

Cloud Computing refers to a comprehensive, expansive system of computing, storage, and networking resources that are easily accessible to the public. 
These resources can be accessed through web service calls over the Internet, allowing for short- or long-term usage based on a pay-per-use model.

\subsection{Virtualization}
Virtualization involves partitioning and sharing hardware resources such as CPU and RAM among multiple virtual machines (VMs). 
A virtual machine monitor (VMM) oversees the allocation of physical resources to running VMs, ensuring performance isolation and security.

Without virtualization, software is closely tied to specific hardware, making it challenging to move or modify applications. 
Failures or crashes are typically isolated to individual servers, operating systems, and applications, resulting in low CPU utilization.

With virtualization, software and hardware become independent, offering greater flexibility with pre-built VMs. 
Operating systems and applications can be managed as unified entities, simplifying deployment and management.

The impact of virtualization on IT systems evolution includes server consolidation and the facilitation of cloud computing.

\paragraph*{Server consolidation}
Server consolidation involves migrating from physical to virtual machines. 
This process allows for seamless movement of virtual machines without disrupting the applications running inside. 
Workloads can be automatically balanced based on predefined limits and guarantees, ensuring efficient resource utilization and protecting servers and applications from component and system failures.
The advantages of server consolidation include:
\begin{itemize}
    \item Running different operating systems on the same hardware, leading to higher hardware utilization.
    \item Reduced hardware requirements, resulting in cost savings in acquisition and management (including human resources, power, and cooling), as well as promoting environmentally friendly practices (Green IT).
    \item Continued use of legacy software, such as running Windows applications on Linux machines through virtualization.
    \item Application independence from underlying hardware, providing flexibility and ease of maintenance.
\end{itemize}