\begin{abstract}
    The lectures are about those topics: 
    \begin{itemize}
        \item Definition of language, theory of formal languages, language operations, regular expressions, regular languages, finite deterministic and non-deterministic automata, 
            BMC and Berry-Sethi algorithms, properties of the families of regular languages, nested lists and regular languages.
        \item Context-free grammars, context-free languages, syntax trees, grammar ambiguity, grammars of regular languages, properties of the families of context-free languages, 
            main syntactic structures and limitations of the context-free languages.
        \item Analysis and recognition (parsing) of phrases, parsing algorithms and automata, push down automata, deterministic languages, bottom-up and recursive top-down syntactic 
            analysis, complexity of recognition.
        \item Translations: syntax-driven, direct, inverse, syntactic. Transducer automata, and syntactic analysis and translation. Definition of  semantics and semantic properties. Static flow analysis of programs. Semantic translation driven by syntax, semantic functions and attribute grammars, one-pass and 
            multiple-pass computation of the attributes.
    \end{itemize}
    The laboratory sessions are about those topics: 
    \begin{itemize}
        \item Modelisation of the lexicon and the syntax of a simple programming language (C-like).
        \item Design of a compiler for translation into an intermediate executable machine language (for a register-based processor).
        \item Use of the automated programming tools Flex and Bison for the construction of syntax-driven lexical and syntactic analyzers and translators.
    \end{itemize}
\end{abstract}