\section{Exercise two}

Why is linear regression important to understand? 
Select all that apply and justify your choice:
\begin{enumerate}
    \item The linear model is often correct.
    \item Linear regression is extensible and can be used to capture nonlinear effects.
    \item Simple methods can outperform more complex ones if the data are noisy.
    \item Understanding simpler methods sheds light on more complex ones.
    \item A fast way of solving them is available.
\end{enumerate}

\subsection{Solution}
\begin{enumerate}
    \item False: it rarely happens that the problem we are modeling has linear characteristics.
    \item True: it is true that this model is easy to interpret and can be extended to also consider nonlinear relationships among variables, e.g., using basis functions.
    \item True: since we are able only to minimize the discrepancy between the considered function and the real one, and we can not reduce the variance introduced by noise, the use of linear model might be a better choice with respect to more complex ones since they are usually prone to overfitting, i.e., they try to model also the noise of the considered process.
    \item True: they are easy to interpret and might give suggestions on more sophisticated techniques which can be used to tackle specific problems.
    \item True/False: for some loss functions we have a closed form solution for linear model (LS method), thus we can guarantee that we are able to find the parameters minimizing the loss function effectively. 
        That is not always true and depends also on the loss function we want to minimize.
\end{enumerate}