\section{CRM}

Multi-channel integration: CRM vision
• Multi-channel integration ensures cross-channel service consistency
• Value extraction from customer data with analytic CRM
• Customer understanding and monitoring with executive functionalities
Operational CRM with different distribution channels: shops/agencies, sales force, call center, Web,
mobile, mail…
Analytical CRM:
• Mining
• Customer profiling
• Campaign management
Executive CRM, reporting
• Service level KPIs
• Customer satisfaction
• Customer profitability
Multi-channel integration ensures cross-channel service consistency
• Value extraction from customer data with analytic CRM
• Customer understanding and monitoring with executive functionalities

\subsection{Analytical CRM}
Mining refers to data analysis with the goal of discovering insights that are relevant
for business management.
• Mining can be performed with any technique: descriptive statistics (e.g. mean
values), data visualization techniques (e.g. plotting data on a bidimensional chart),
statistical correlations (e.g. correlating IT investments with sales). More recently,
also with machine learning techniques.
• Why mining data?
– Several companies have huge operational databases embedding new knowledge (KM
information process)
– Data mining is used to extract patterns
– Patterns should be explained by summarizing findings in an intuitive way (e.g. simple statistical
indicators, recurring associations, etc.)
Customers who can be traced (with a loyalty card) can be segmented according
to different dimensions:
- Loyalty, from the analysis of their purchasing habits, such as frequency,
recency and expenditure.
- Price sensitivity, such as up, low and mid market.
- Lifestyle, that is the behavioural orientation of customers along different
dimensions, such as «business», «casual», «classic», «vintage» or «sports
wear»
Customer segmentation represents the basis for targeting promotions, by:
- Targeting a segment or a mix of segments.
- Using the average characteristics of a segment to complement the
knowledge on individual customers
Campaign management
Campaign management is a set of functionalities that support marketing
campaigns. Campaigns have four fundamental phases and related functionalities:
Phases Planning and 
budgeting Design Execution Evaluation 

\subsection{Operational CRM}
Sales Force Automation (SFA)
• Goal: to provide CRM front-end functionalities to the sales force in both B2C and B2B
• Advantages for companies:
– Management perspective: governance of the sales force.
– Sales force perspective: stronger sales capabilities.
• Operating objectives :
– Reduce the costs of customer acquisition.
– Increase customer retention.
– Reduce bureacracy for customers and increase responsiveness during the sales process.
– Increase the effectiveness of the sales force
Sales process on physical channels
CRM provides the data that allow the definition of individual incentive
systems
• The sales force must have access to all information on customers,
including invoicing.
Physical channel: performance indicators (KPIs)
• Effectiveness
– Overall effectiveness= Contracts/Opportunities
– Effectiveness of contacts = Contacts /Contracts
– Investment in contacts= Contacts/Customers
• Efficiency
– Productivity= Sales (in €)/ number of sales people
– Unit costs= Cost of sales force (in €) / Contracts (number)
• Service level
– Customer satisfaction = Value/Price.
– Churn = Lost clients (number or sales in € or life-time value) / Total number of clients (number or
sales in € or life-time value)
– Retention = Loyal clients (number or sales in € or life-time value) /Total number of clients
(number or sales in € or life-time value)
The physical channel is expensive, it is suitable for «rich» products, especially B2B
• The sales force tends to become the «owner» of data on customers
• Pharmaceutical industry in the US
• 90.000 promoters of drugs
• 18 million presentations of drugs/year
• 650.000 physicians.
• Each promoter is responsible for \$ 1,5 million.
• Insurance
– Agencies and agents.
– Agents have an end-to-end responsibility of their customers, from contact to post-sale and invoicing.
• Finance:
– Agents are more and more focused on private banking and consulting (value-added services)
• Telephone companies
– Agents are limited to B2B
– Service is custom designed in B2B
– Document management
A document is composed by:
• Static sections (logo, address, …)
• Dynamic sections updated by the
configuration and order management tools.

