\section{Grammar normal forms and transformations}

Grammars normal forms constrain the rules without reducing the family of generated languages. 
They are useful for both proving properties and for language design. 

Let us see some transformations useful both to obtain an equivalent normal form and design the syntax analyzers.

\subsection*{Nonterminal expansion}
The expansion of a nonterminal is used to eliminate it from the rules where it appears. 
\begin{example}
    Consider the grammar: 
    \[
    \begin{cases}
        A \rightarrow \alpha B \gamma \\
        B \rightarrow \beta_1|\beta_2|\dots|\beta_n
    \end{cases}    
    \]
    With the expansion of the nonterminal $B$ we obtain: 
    \[A \rightarrow \alpha\beta_1\gamma|\alpha\beta_2\gamma|\dots|\alpha\beta_n\gamma\]
\end{example}

\subsection*{Elimination of the axiom from right parts}
It is always possible to obtain right part of rules as strings by simply introducing a new axiom $S_0$ and the rule $S_0 \rightarrow S$.

\subsection*{Normal form without nullable nonterminals}
A non-terminal A is nullable if it can derive the empty string. 

Consider the set Null $\subseteq V$ of nullable non-terminals. 
It is composed of the following logical clauses, to be applied until a fixed point is reached:
\[A \in \textnormal{Null} \implies
\begin{cases}
    (A \rightarrow \varepsilon) \in P\\
    (A \rightarrow A_1A_2\dots A_n) \in P \:\:\:\:\:\: \textnormal{with } A_i \in V \backslash \{A\} \\
    \forall 1 \leq i \leq n \:\:\:\:\:\: \textnormal{with } A_i \in \textnormal{Null}
\end{cases}    
\]
The construction of the non-nullable form consist in:
\begin{enumerate}
    \item Compute the Null set. 
    \item For each rule within $P$ add as alternatives those obtained by deleting, in the right part, the nullable non-terminals. 
    \item Remove all empty rules, except for $A=S$. 
    \item Clean the grammar and remove any circularity. 
\end{enumerate}
The normal form without nullable nonterminals needs that no nonterminal other than the axiom is nullable. 
In that case the axiom is nullable only if the empty string $\varepsilon$ is in the language. 












