\section{Support Vector Machines}

Support Vector Machines are a flexible and theoretically supported method initially applied to classification. 
Over the years, they have been extended to handle regression, clustering, and anomaly detection problems.

The fundamental idea behind SVM is to find the hyperplane that maximizes the margins (distance between the boundary and the points).

The hypothesis space of SVM is defined as $y(\mathbf{x})=f(\mathbf{x},\mathbf{w})=\text{sign}(\mathbf{w}^T\mathbf{x}+b)$. 
The loss function is computed over the dataset $\mathcal{D}=\left\{(\mathbf{x}_n,t_n)\right\}_{n=1}^{N}$ with $t_n\in\{-1,1\}$.
It aims to minimize:
\[\dfrac{1}{2}\left\lVert \mathbf{w} \right\rVert _2^2+C\sum_{n=1}^{N}\zeta_i \]
with respect to $\mathbf{w},\zeta_1,\ldots,\zeta_N$ subject to:
\[\begin{cases}
    t_n(\mathbf{w}^T\mathbf{x}_n+b)\geq 1-\zeta_i \qquad\forall n \in \{1,\ldots,N\} \\
    \zeta_i\geq 0\qquad\forall\in\{1,\ldots,N\}
\end{cases}\]
where $C > 0$ is a hyperparameter. 
The optimization method typically used for SVM is sequential quadratic optimization.

\paragraph*{Python}
To train a linear classification Support Vector Machine: 
\begin{enumerate}
    \item Define the SVM: \texttt{SVM\_model.svm.SVC(kernel='linear')}.
    \item Train the SVM: \texttt{SVM\_model.fit(input, target)}.
\end{enumerate}
We are interested in determining:
\begin{itemize}
    \item Boundary $\textbf{w}^T\textbf{x}_n+b=0$.
    \item Margins $\textbf{w}^T\textbf{x}_n+b=\pm 1$.
    \item Support vectors (\texttt{SVM\_model.support\_vectors\_}).
\end{itemize}
The use of kernels in SVMs is almost native, turning it into a non-parametric method:
\begin{itemize}
    \item Hypothesis space: $y(\mathbf{x})=f(\mathbf{x},\mathbf{w})=\text{sign}\left(\sum_{n=1}^{N}\alpha_nt_nK(\mathbf{x}_n,\mathbf{x})+b\right)$. 
    \item Loss measure: loss function in the dual formulation.
    \item Optimization method: quadratic optimization.
\end{itemize}
In Python:
\begin{itemize}
    \item Define the SVM: \texttt{SVM\_model = svm.SVC()}.
    \item Train the SVM: \texttt{SVM\_model.fit(input, target)}.
\end{itemize}
With kernels, we no longer have an explicit formula for the boundary and the margins.