\section{Design principles}

Effective mobile app design revolves around several core principles that prioritize usability, clarity, and user engagement. These principles guide the creation of intuitive and enjoyable user experiences:
\begin{enumerate} 
    \item \textit{Keep it brief}: present information concisely to avoid overwhelming the user. 
    \item \textit{Pictures are faster than words}: visual elements communicate more efficiently than text, improving comprehension and engagement.
    \item \textit{Decide for me but let me have the final say}: offer smart defaults and automation to simplify tasks, while still allowing users to override settings if needed. 
    \item \textit{I should always know where I am}: ensure users have a clear sense of navigation and context within the app, reducing confusion and frustration. 
\end{enumerate}

To successfully implement these principles, designers should adopt the following approaches:
\begin{itemize} 
    \item \textit{Mobile mindset}: design with a mobile-first approach, focusing on creating experiences that are streamlined, unique, and user-centered. 
        Apps should be functional, visually appealing, and tailored to mobile usage patterns.
    \item \textit{Identify different classes of users}: clearly define and understand the needs of your target audience. 
        Users can be categorized into groups such as those who are bored, busy, or lost, each with distinct behaviors and expectations.
    \item \textit{First Impressions Matter}: capture user interest within the first few seconds of interaction. 
        To make a lasting impact:
        \begin{itemize}
            \item Provide minimal or no help text, relying instead on intuitive design.
            \item Create a distinctive and captivating look and feel.
            \item Ensure the app quickly conveys its purpose and value.
        \end{itemize}
\end{itemize}