\section{Exercise one}

A computer system is engineered with a failure rate of one fault every five years under typical usage conditions. 
This system lacks fault tolerance capabilities, meaning it ceases functioning upon encountering its initial fault.
\begin{enumerate}
    \item Compute the mean time to failure of the system.
    \item Compute the probability that the system will fail during its first year of operation. 
    \item The standard warranty for the system covers 2 years of operation. 
        However, the vendor aims to provide extended insurance against failures for the initial 5 years of operation, for which they plan to charge an additional fee.  
        The vendor intends to charge $20\$$ for every $1\%$ decrease in reliability to offer this insurance. 
        Compute how much should the vendor charge for such an insurance.
\end{enumerate}

\subsection*{Solution}
\begin{enumerate}
    \item The mean time to failure for the system is calculated as the inverse of the failure rate:
        \[\text{MTTF}=\dfrac{1}{\lambda}=\dfrac{1}{\frac{1}{5}}=5\text{ years}\]
    \item The failure probability within a given time frame is given by:
        \[F(t)=1-R(t)\]
        For the first year:
        \[F(1)=1-R(1)=1-e^{-\lambda}=1-0.818=0.18\]
    \item We compute the reliability at 2 and 5 years:
        \[R(2\text{ year})=e^{-2\lambda}=0.67\]
        \[R(5\text{ year})=e^{-2\lambda}=0.37\]
        This yields a reliability drop of $0.3$. 
        Since the vendor wants to charge $20\$$ for each $1\%$ drop in reliability, the cost is:
        \[\text{cost}=\dfrac{0.3}{0.01}=30\cdot 20\$=600\$\]
\end{enumerate}