\section{Introduction}

React Native is an open-source framework derived from React, tailored specifically for mobile application development. 
It allows developers to build cross-platform applications for Android and iOS using a single codebase while leveraging each platform's native capabilities. 
By combining JavaScript with native APIs, React Native simplifies mobile development without sacrificing the performance and look of fully native applications.
React is used to define the UI components and their behavior, making it a powerful tool for creating responsive and dynamic mobile user interfaces.

While React Native is predominantly associated with Android and iOS development, its reach extends far beyond these platforms. 
The framework has evolved to support:
\begin{itemize}
    \item \textit{React Native for Windows and macOS}: enabling native app development for desktop environments.
    \item \textit{React Native for tvOS}: targeting Apple TV and Android TV applications.
    \item \textit{React Native Web}: allowing React Native apps to run on the web through React DOM.
    \item \textit{React Native Skia}: a renderer that supports rendering on Linux and macOS using the Skia graphics engine.
\end{itemize}

At the core of React Native are a set of pre-built native (core) components, covering a wide range of functionality from UI controls to activity indicators. 
These components are ready to use out-of-the-box, streamlining development processes. 
Additionally, React Native provides flexibility for developers to create their own custom native components for Android and iOS, allowing for fine-tuned customization to meet specific project needs.

Both React and React Native employ JSX, a syntactic extension that allows developers to write UI components seamlessly within JavaScript, offering a clean, declarative approach to building user interfaces.
This combination of cross-platform compatibility, customizable components, and modern development practices makes React Native a powerful and versatile framework for building high-quality mobile applications.