\section{Exercise one}

\begin{enumerate}
    \item Show that the $VC$ dimension of an axis aligned rectangle is 4.
    \item Show that the $VC$ dimension of a linear classifier in 2D is 3.
    \item Show that the $VC$ dimension of a triangle in the plane is at least 7.
    \item Show that the $VC$ dimension of a 2D stump, i.e., use either a single horizontal or a single vertical line in 2D to separate points in a plane, is 3.
\end{enumerate}

\subsection*{Solution}
\begin{enumerate}
    \item Consider 4 points. 
        It is possible to show by enumeration that all the possible labeling are shattered by the rectangle. 
        Consider 5 points. 
        Consider the set of points with maximum and minimum $x$ coordinate and maximum and minimum $y$ coordinates. 
        If all the points are on the rectangle, we consider the labeling which assign alternate labels to the points if you follow the rectangle perimeter. 
        Otherwise, there are at most 4 points in this set. 
        If we label them $+$ and label $-$ the other, it is not possible to shatter this labeling.
    \item Let us call the class of all the linear classifier in 2D $\mathcal{H}$. 
        The proof consists in two steps: 
        \begin{enumerate}
            \item $V C(\mathcal{H}) \geq 3$ and $V C(\mathcal{H}) \leq 3$. 
                We need to show that it exists a set of 3 points that can be shattered by the considered hypothesis space $\mathcal{H}$. 
                By considering a set of three non-aligned points, it is possible to show by enumeration that it is possible to shatter them with a linear classifier. 
                Thus, $V C(\mathcal{H}) \geq 3$. 
            \item We need to show that it does not exists a set of 4 points which can be shattered by a linear classifier. 
                There are different cases:
                \begin{enumerate}
                    \item Four aligned points: if we alternate instances coming from the positive and negative classes we cannot shatter them.
                    \item Three aligned points and a fourth on an arbitrary position: if we alternate instances coming from the positive and negative classes for the three aligned points, we cannot shatter them.
                    \item Four points on a convex hull: if we label the points on the two diagonals with opposite classes, we cannot shatter them.
                    \item Three points on a convex hull (triangle) and one inside the hull: if we label the three points on the triangle with a label and the last one with the other class, we cannot shatter them.
                \end{enumerate}
                Since there does not exists a configuration where we can shatter the points, we have that $V C(\mathcal{H}) \leq 3$.          
        \end{enumerate}
    \item If we consider a set of points on a circle it is possible to show by enumeration that a triangle ia able to shatter all of them.
    \item Since the decision stumps in 2D are a model which is less flexible than linear boundaries, which have $V C = 3$, they should have $V C(\mathcal{H}) \leq 3$. 
        The proof that $V C(\mathcal{H}) = 3$ is by enumeration:
\end{enumerate}