\section{Formal language theory}

A formal language consists of words formed by selecting letters from an alphabet, and these words must adhere to a defined set of rules to be considered well-structured.
\begin{definition}[\textit{Alphabet}]
    An alphabet, denoted by $\Sigma$, is a finite set of elements referred to as characters and represented as $\{a_1,a_2,\dots,a_k\}$. 
\end{definition}
\begin{definition}[\textit{Alphabet cardinality}]
    The alphabet cardinality, denoted as $\left\lvert\Sigma\right\rvert$, is the number of characters in a finite alphabet $\Sigma$, expressed as $\left\lvert\Sigma\right\rvert=k$.
\end{definition}
\begin{definition}[\textit{String}]
    A string is a sequential arrangement of elements from an alphabet, potentially containing repetitions.
\end{definition}
\begin{example}
    The alphabet $\Sigma={a,b}$ comprises two distinct characters.
    This alphabet allows for the generation of various languages, including:
    \begin{itemize}
        \item $L_1=\{aa,aaa\}$
        \item $L_2=\{aba,aab\}$
        \item $L_3=\{ab,ba,aabb,abab,\dots,aaabbb,\dots\}$
    \end{itemize}
    In these languages, different combinations of the alphabet's characters are used to construct words.
\end{example}
\begin{definition}[\textit{Sentence}]
    The strings of a language are called sentences or phrases. 
\end{definition}
\begin{definition}[\textit{Language cardinality}]
    The cardinality of a language is the total number of sentences it contains.
\end{definition}
\begin{example}
    Consider the language $L_2={ bc, bbc }$, which consists of two elements, indicating a cardinality of two.
\end{example}
\begin{definition}[\textit{Number of occourences}]
    The number of occurrences is the count of times a specific letter appears in a word.
\end{definition}  
\begin{definition}[\textit{String length}]
    The length of a string is the total number of elements it contains.
\end{definition}
\begin{definition}[\textit{String equality}]
    Two strings are considered equal if and only if they have the same length, and their elements match from left to right, sequentially.
\end{definition}
\begin{example}
    For the string $aab$, the counts of occurrences for the letters $a$ and $c$ are represented as follows:
    \[{\left\lvert aab \right\rvert}_a = 2\]
    \[{\left\lvert aab \right\rvert}_c = 0\]
    The length of the string $aab$ is calculated as:
    \[\left\lvert aab \right\rvert = 3\]
\end{example}