\section{Graph theory}

A graph is a mathematical structure consisting of nodes (or vertices) connected by edges (or lines).
\begin{definition}
    Let $G$ be defined as the ordered triple $G=(V, E, f)$, where:
\end{definition}
\begin{itemize}
    \item $V$ represents a set of nodes, points, or vertices. 
    \item $E$ is a set of edges, which are the connections between the nodes.
    \item $f$ is a function that associates each edge in $E$ with an unordered pair of vertices in $V$. 
\end{itemize}
\begin{definition}[\textit{Path}]
    A path is defined as a sequence of vertices such that there exists an edge leading from each vertex to its successor. 
\end{definition}
A path is considered simple if all vertices are distinct. 
A path that starts and ends at the same vertex is known as a cycle.
\begin{definition}[\textit{Cyclic graph}]
    A graph is termed cyclic if it contains at least one cycle.
\end{definition}
\begin{definition}[\textit{Connected}]
    A graph is connected if there exists a path between any pair of vertices, allowing traversal from one node to any other. 
\end{definition}
\begin{definition}[\textit{Directed graph}]
    A directed graph, or digraph, is characterized by edges that have a direction.
\end{definition}
\begin{definition}[\textit{Strongly connected}]
    A directed graph is strongly connected if there is a directed path from any node to any other node.
\end{definition}
\begin{definition}[\textit{Tree}]
    A tree is a special type of graph that is connected and acyclic, where there is exactly one path between any two nodes.
\end{definition}

\subsection{Graph data structure}
In computer science, a graph is defined as an abstract data type composed of a set of nodes, and a set of edges. 
These elements establish relationships between the nodes, reflecting the mathematical concept of graphs.
Graphs can be represented in several ways, including:
\begin{itemize}
    \item \textit{Matrix representation}: incidence matrix (edge data in relation to vertices, size $|E|\times|V|$) or adjacency matrix (adjacency or edge weights, size $|V|\times|V|$). 
    \item \textit{List representation}: edge list (pairs of vertices with optional weights and additional data), and adjacency list (collection of lists or arrays where each list corresponds to a vertex and contains its adjacent vertices).
\end{itemize}