\section{Exercise one}

As you're fitting a linear model to your dataset, you contemplate transitioning to a quadratic model, incorporating quadratic features $\varphi(x) =\begin{bmatrix} 1 & x & x^2 \end{bmatrix}$. 
Considering this change, which of the following statements is most likely true:
\begin{enumerate}
    \item Employing the quadratic model will reduce your reducible error.
    \item Employing the quadratic model will decrease the bias in your model.
    \item Employing the quadratic model will reduce the variance in your model.
    \item Employing the quadratic model will decrease your reducible error.
\end{enumerate}
Provide motivations to your answers.

\subsection*{Solution}
\begin{enumerate}
    \item False: changing the model doesn't directly impact the irreducible error, as it's inherent to the problem itself and can't be mitigated by model choice.
    \item True: expanding the model to include quadratic features increases its flexibility, making it better able to capture complex relationships in the data. 
        Therefore, it's likely to reduce bias or, at the very least, not increase it.
    \item False: introducing more complex features typically leads to increased variance as the model becomes more sensitive to fluctuations in the training data.
    \item Partially true: whether using the quadratic model decreases the reducible error depends on the balance between bias reduction and variance increase. 
        If the quadratic model effectively reduces bias without excessively inflating variance, it can lead to a more accurate model and decrease reducible error. 
        However, if the increase in variance outweighs the reduction in bias, the overall error may increase. 
        Therefore, the statement could be true or false depending on the specific circumstances.
\end{enumerate}