\section{Formal language theory}

A \emph{formal language} consists of words whose letters are taken from an alphabet and are well-formed according to a specific set of rules.
\begin{definition}
    An \emph{alphabet} is a finite set of elements called terminal symbols or \emph{characters}. 
    The \emph{cardinality} of an alphabet \[\Sigma =\{ a_1,a_2,\dots, a_k \}\] is the number of characters that it contains: $\left\lvert \Sigma \right\rvert = k$. 
    A \emph{string} or word is a sequence of characters. 
\end{definition}
\begin{example}
    The alphabet $\Sigma =\{ a,b \}$ has a cardinality of two. Some possible languages derived from this alphabet can be:
    \begin{itemize}
        \item $L_1=\{aa,aaa\}$
        \item $L_2=\{aba,aab\}$
        \item $L_3=\{ab,ba,aabb,abab,\dots,aaabbb,\dots\}$
    \end{itemize}
\end{example}
\begin{definition}
    Given a language, a string belonging to it is called a \emph{sentence} or \emph{phrase}. The \emph{cardinality} or size of a language is the number of sentence it contains.
    If the cardinality is finite, the language is called \emph{vocabulary}. 
\end{definition}
\begin{example}
    Given the language (that is a vocabulary) $L_2=\{ bc,bbc \}$ we have that its cardinality is equal to two. 
\end{example}
\begin{definition}
    The number of repetitions of a certain letter in a word is called \emph{number of occurrences}. The \emph{length} of a string is the number of its elements. 
    Two strings are \emph{equal} if and only if: 
    \begin{itemize}
        \item They have the same length.
        \item Their elements, from left to right, coincide. 
    \end{itemize}
\end{definition}
\begin{example}
    The number of occurrences of $a$ and $c$ in $aab$ is indicated with:
    \[{\left\lvert aab \right\rvert}_a = 2\]
    \[{\left\lvert aab \right\rvert}_c = 0\]
    The length of the string $aab$ is equal to: 
    \[\left\lvert aab \right\rvert = 3\]
\end{example}